---
title: "Admission Evaluation Criteria Analysis"
subtitle: "DSAN 5900: Digital Storytelling"
authors:
  - name: Bella Shi
    email: js5081@georgetown.edu
    affiliation: Georgetown University
    corresponding: true  
  - name: Samantha Moon
    email: sjm206@georgetown.edu 
    affiliation: Georgetown University
    corresponding: true
  - name: Lianghui Yi
    email: ly297@georgetown.edu
    affiliation: Georgetown University
    corresponding: true    
df-print: kable
title-block-banner: "#E69F0095"
title-block-banner-color: "black"
code-fold: true
format:
  html:
    df-print: kable
  pdf:
    link-citations: true 
prefer-html: true
---

# Introduction
This report provides an analysis of the admissions data to identify key factors influencing admissions decisions. The goal is to generate actionable insights for institutional leadership to support strategic decision-making.


## Objectives
- Understand the distribution of admissions decisions across states.
- Analyze the impact of GPA, test scores, work experience, and volunteer levels on admissions.
- Provide actionable recommendations based on data-driven insights.

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(dplyr)
library(ggcorrplot)

# Load the data
data <- read_csv('SummerStudentAdmissions2.csv')

```

```{r}
# Initial data check
head(data)
```


```{r}
summary(data)
```

## Data Cleaning
```{r data-cleaning}
# Remove invalid 'Decision' values
valid_decisions <- c('Admit', 'Decline', 'Waitlist')
data <- data %>% filter(Decision %in% valid_decisions)

# Replace invalid gender values (-1.0) with NA
data$Gender <- ifelse(data$Gender == -1, NA, data$Gender)

# Handle missing values
data <- data %>% drop_na()

# Check cleaned data
summary(data)
```

## Visualizations and Analysis


### 1. Admission Rates by State

```{r admission-rates-by-state}
ggplot(data, aes(x = State, fill = Decision)) +
  geom_bar(position = 'fill') +
  labs(title = 'Admission Rates by State', y = 'Proportion', x = 'State') +
  theme_minimal()
```
**Analysis:** The visualization highlights which states have higher or lower admission rates. States with consistently low acceptance rates may need targeted recruitment strategies or support programs.



### 2. GPA and Test Scores for Admitted vs. Rejected Students

```{r gpa-testscore-distribution}
ggplot(data, aes(x = GPA, y = TestScore, color = Decision)) +
  geom_point() +
  labs(title = 'GPA vs. Test Scores by Admission Decision', x = 'GPA', y = 'Test Score') +
  theme_minimal()
```
**Analysis:** This scatter plot reveals how GPA and test scores correlate with admissions decisions. Students with higher GPAs and test scores generally have a higher chance of being admitted.



### 3. Correlation Heatmap

```{r correlation-heatmap}
corr_matrix <- data %>% select(GPA, WorkExp, TestScore, WritingScore, VolunteerLevel) %>% cor()
ggcorrplot(corr_matrix, lab = TRUE, title = 'Correlation Matrix')
```
**Analysis:** The heatmap shows the correlations between quantitative variables. Strong correlations can help identify which metrics are most predictive of admissions success.



## Conclusion

The analysis indicates that GPA, test scores, and volunteer experience significantly influence admissions decisions. The state-wise analysis suggests opportunities for targeted recruitment. To enhance admission strategies, the institution could:

- Focus on supporting applicants from states with lower acceptance rates.
- Prioritize academic and extracurricular excellence in admissions criteria.
- Consider refining recruitment efforts based on insights gained from the data.
